# Getting and Cleaning Data Course Project

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Download the data file and unzip it to "project" folder.


```{r}
library(data.table)
library(plyr)

if( !file.exists("./Data") ) { dir.create("./Data") }

if( !file.exists("UCI_HAR_Dataset.zip")) {
      fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
      download.file(fileUrl, destfile="./UCI_HAR_Dataset.zip")
}

unzip(zipfile="./UCI_HAR_Dataset.zip", exdir = "./Data")
dataPath <- file.path("./Data", "UCI HAR Dataset")

```

# 2.Merge the training and the test sets to create one data set.
## 2.1 Read the data files 

```{r}
# read the train and test data set, each row is a 561-element vector 
xTrain <- data.frame(fread(file.path(dataPath, "train", "X_train.txt")))
xTest<- data.frame(fread(file.path(dataPath, "test", "X_test.txt")))

#read the subject data, the data range of is from 1 to 30
subjectTrain <- data.frame(fread(file.path(dataPath , "train", "subject_train.txt")))
subjectTest <- data.frame(fread(file.path(dataPath , "test" , "subject_test.txt" )))

#read the activity data, the data range is from 1 to 6
activityTrain <- data.frame(fread(file.path(dataPath , "train", "Y_train.txt")))
activityTest  <-data.frame( fread(file.path(dataPath , "test" , "Y_test.txt" )))

#read feature names
featureNames <- fread(file.path(dataPath, "features.txt"))


```

## 2.2 Merge train and test data

```{r}
subjectData <- rbind(subjectTrain, subjectTest)
activityData <- rbind(activityTrain, activityTest)
featureData <- rbind(xTrain, xTest)

```
## 2.3 Set column names 

```{r}
setnames(subjectData, "V1", "subject")
setnames(activityData, "V1", "activity")

#Appropriately labels the data set with descriptive variable names.
setnames(featureData, names(featureData), featureNames$V2)



```
## 2.4 Merge the subject data, activity data and feature data into one big data set

```{r}
mergedData <- cbind(cbind(subjectData, activityData), featureData)
str(mergedData)


```
# 3. Extracts only the measurements on the mean and standard deviation for each measurement

```{r}
#extract features name with "mean()" or "std()"
selectedFeature <- featureNames$V2[grep("mean\\(\\)|std\\(\\)", featureNames$V2)]
selectedFeature
mergedData <- mergedData[,c("subject", "activity", selectedFeature[1:length(selectedFeature)])]
str(mergedData)

```
# 4. Uses descriptive activity names to name the activities in the data set

```{r}
activityNames <- fread(file.path(dataPath, "activity_labels.txt"))
mergedData$activity <- factor(mergedData$activity, levels=activityNames$V1, labels=activityNames$V2) 
head(mergedData$activity, 30)

```
#5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
## 5.1 Create a second, independent tidy data set
```{r}
#Group the data by subject and activity, and get the mean of each other variables
tidyDataSet <- ddply(mergedData, .(subject, activity), colwise(mean))
str(tidyDataSet)

```
## 5.2 Write the second tidy data set to txt file
```{r}
write.table(tidyDataSet, file = "tidyDataSet.txt", row.names = FALSE)

```
